### 2016-05-29 thesis_paper/source.tex
paper stuff
paper

\subsubsection{Two-body functions implementation in GPU Version} Once the first kernel is finished we run the two-body kernel with its own stream. As it has already been mentioned, the biggest workload is in this kernel, because it has higher complexity. In this particular case it is $O(n^2)$. GPU streaming multiprocessors are very powerful for straightforward primitive independent calculations, and each core is much weaker than a moden CPU which has a whole bunch of L caches with some prediction behavior systems for code divergence. As you can see on Figure \ref{fg:threadload}, unbalanced thread load assumes for GPU kernel to check every time if the target atom can interact with other atom and if it has to wait till other threads finish their jobs. Thus, we had to come up with a strategy of making equally balanced thread jobs and distribute them among the threads leaving atom to thread mapping as is.

For SDH we the following strategy

\begin{itemize}
\item[1.] Find current index (for target atom)
\item[2.] Find range of atoms it has to interact with in an equally loaded fashion
\item[3.] Load the range of atoms into shared memory
\item[4.] Iterate through the range of atoms from shared memory and atomically write the distances counts into buckets pre-allocated in share memory
\item[5.] Synchronize with other threads in the block (not in the entire grid)
\item[6.] If it is the first thread in block, add all bucket counts from shared memory into global memory
\end{itemize}

>>>>

\subsubsection{Two-body functions implementation in GPU Version}

The approach of solving SDH with GPU is absolutely different just like with one-body queries, though number of computations are the same. We do brute force  $O(\frac{N^2}{2})$ complex solution which is pretty much looking at distance between every pair of atoms. In Pseudocode \ref{alg:sdhgpu} we have tried to provide main logic flow that explains how exactly we implemented kernel for SDH computation. 
Once the first kernel is finished we run the two-body kernel with its own stream. As it has already been mentioned, the bigger workload is in this kernel, because it has higher complexity. GPU streaming multiprocessors are very powerful for straightforward primitive independent calculations, and each core is much weaker than a moden CPU which has a whole bunch of L caches with some prediction behavior systems for code divergence. As you can see on Figure \ref{fg:threadload}, unbalanced thread load assumes for GPU kernel to check every time if the target atom can interact with other atom and if it has to wait till other threads finish their jobs. Thus, we had to come up with a strategy of making equally balanced thread jobs and distribute them among the threads leaving atom to thread mapping as is.

Here are the main steps of the algorithm:
\begin{itemize}
\item[1.] Get mapped atom
\item[2.] Load all block related atoms to shared memory
\item[3.] Go pairwise with atoms for SDH in a equal balance manner
\item[4.] Synchronize threads
\item[5.] Get everything block shared memory to global memory
\end{itemize}

\begin{figure}
 \centerline{ \includegraphics[width=0.5\columnwidth]{images/sharedMem}}
 \caption{Representation of example of shared memory usage}
 \label{fg:sharedMem}
\end{figure}

As we already mentioned, the main purpose of our proposal is the system that allows integration of GPU into push-based system, thus, we didn't go with too sophisticated and optimized implementations for each query that we benchmark, assuming that user's wouldn't do it and having more or less realistic speedup results. Figure \ref{fg:sharedMem} shows swimlanes as execution thread for blocks. Every row is an execution block for an atom, where first element is the mapped atom. First of all, you can observe that we can clearly see that the execution threads are almost equally balanced (exceptions on odd number of atoms). Every swimlane (bordered with black) represent within block execution. Colored numbers are the atoms that we loaded into shared memory. Of course, this is about half of improvement, and we can happily add improvement of adding second part too, but, at this point, we go with this average improvement approach.

%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%      SDH GPU      %%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{algorithm}
\caption{Kernel function implementation for SDH in GPU}
\label{alg:sdhgpu}
\begin{algorithmic}[1]
\Procedure{gpusdhkernel}{atomsCount, pdhRes, histogram, atoms, numOfBuckets}
\State \Comment{ $atoms$ is vector of atoms kept in global gpu memory}
\State \Comment{ $atomsCount$ is number of atoms in the frame kept in global gpu memory}
\State \Comment{ $histogram$ is buckets with counters for the SDH}
\State \Comment{ $pdhRes$ is resolution of histogram}
\State \Comment{ $numOfBuckets$ is number of buckets}

\State $\textit{smem} \gets \textit{allocate shared memory in each GPU block for fast write access of histogram buckets}$

\State $\textit{sharedHisto} \gets \textit{pointer to the head of smem}$
\State $\textit{sharedAtoms} \gets \textit{pointer to the tail of sharedHisto}$
\State $\textit{indexX and indexY} \gets \textit{calculate from grid sizes}$ 
\State $\textit{index} \gets \textit{get atom mapped to this kernel}$
\State \Comment{Finally get the index of current atom mapped to this kernel}

\If {$\textit{threadIdx.x} = 0$}

\State $\textit{sharedHisto} \gets \textit{Set all values to 0}$

\State $\textit{start} \gets \textit{index}$
\State $\textit{k} \gets \textit{0}$
\State $\textit{i} \gets \textit{index}$
\For {$i < start + blockDim.x$ and $i < atomsCount$}

	\State $\textit{sharedAtoms[k]} \gets \textit{atoms[i]}$
	\Comment{Load local atoms into shared memory}

	\State $\textit{k} \gets \textit{k + 1}$
	\State $\textit{i} \gets \textit{i + 1}$
\EndFor

\EndIf

\State
\State $\textit{\_\_syncthreads()}$
\State $\textit{threadLoad} = \textit{(atomsCount + 1) / 2}$
\State $\textit{start} \gets \textit{i+1}$
\State $\textit{end} \gets \textit{i + threadLoad}$

\State $\textit{end} \gets \textit{increment by one if it is odd and total count is even}$

\State $\textit{bi} \gets \textit{blockDim.x * blockIdx.x}$
\State $\textit{ei} \gets \textit{bi + blockDim.x}$

\State $\textit{ind1} \gets \textit{threadIdx.x}$

\State $\textit{j} \gets \textit{start}$

\For {$\textit{j} < \textit{end}$}

\State $\textit{ind2} \gets \textit{j mod atomsCount}$
\State $\textit{atom1} \gets \textit{load from shared memory}$

\If{ $\textit{atom2 can be loaded from shared memory}$}
\State $\textit{atom2} \gets \textit{load from local memory}$
\Else
\State $\textit{atom2} \gets \textit{load from global memory}$
\EndIf

\State $\textit{calc distance and atomically write to local histogram}$

\EndFor
\State
\State $\textit{synchronize all threads}$
\State $\textit{gather all histogram from shared into global memory}$
\EndProcedure
